---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(stringr)
library(ggplot2)
```

## Loading and preprocessing the data
```{r}
repData <- read.csv("./activity.csv") ##Set the directory where data are stored                                          as working directory.
repData$date <- as.Date(repData$date)
repData$steps <- as.numeric(repData$steps)
repData$interval <- as.numeric(repData$interval)
str(repData)
##repData$date <- as.Date(as.character(repData$date), format = "%Y/%m/%d")
```
## What is mean total number of steps taken per day?
```{r}
repData2 <- repData[which(!is.na(repData$steps)),]
day.sum <- tapply(repData2$steps, repData2$date, sum)
hist(day.sum, 10, main ="Total steps per day")
```

The mean steps taken per day is `r mean(day.sum)` steps and the median amount of steps taken per day is `r median(day.sum)` steps.

## What is the average daily activity pattern?
```{r}
interval.activity <- repData2 %>% group_by(interval) %>% summarize(avg_steps = mean(steps))
ggplot(data = interval.activity, aes(interval, avg_steps)) + geom_line()

print(max(interval.activity$avg_steps))
```
The interval `r interval.activity[interval.activity$avg_steps==max(interval.activity$avg_steps),1]` has the highest average amount of steps. There is on average `r max(interval.activity$avg_steps)` steps for this interval.  

## Imputing missing values
```{r}
dailyact <- tapply(repData2$steps, repData2$interval, mean)
missing <- sum(is.na(repData))
repData3 <- repData
repData3[which(is.na(repData3$steps)),1] <-  dailyact[as.character(repData3[which(is.na(repData3$steps)),3])]
sum(is.na(repData3))


new.sum <- tapply(repData3$steps, repData3$date, sum)
hist(day.sum,10, main = "Total number of steps taken per day", xlab = "Steps"
     , ylim =c(0, 25))
abline(v = median(new.sum), col = 4, lwd = 4)
hist(new.sum,10, main = "Total number of steps taken per day  
     (missing values replaced with mean of interval)", xlab = "Steps",
     ylim =c(0, 25))
abline(v = median(new.sum), col = 4, lwd = 4)

mean(new.sum)
median(new.sum)
mean(new.sum) - mean(day.sum)
median(new.sum) - median(day.sum)

```

There are `r sum(is.na(repData))` observations missing. The mean steps taken per day, with the NA values replaced with the interval average, is `r mean(new.sum)` steps and the median steps taken is `r median(new.sum)`. There is no difference between the average steps taken per day when NA's are replaced, and about a one step difference in the median.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
repData3$wd<-weekdays(repData3$date)
repData3$fwd<- as.factor(c("weekend", "weekday"))
repData3[repData3$wd == "Sunday" | repData3$wd == "Saturday" ,5]<- factor("weekend")
repData3[!(repData3$wd == "Sunday" | repData3$wd == "Saturday"),5 ]<- factor("weekday")

repData3_we <- subset(repData3, fwd == "weekend") 
repData3_wd <- subset(repData3, fwd == "weekday") 
dailyact_we<-tapply(repData3_we$steps, repData3_we$interval, mean)
dailyact_wd<-tapply(repData3_wd$steps, repData3_wd$interval, mean)
par(mfrow=c(2,1))
plot(y = dailyact_wd, x = names(dailyact_wd), type = "l", xlab = "5-Minute Interval", 
     main = "Daily Activity Pattern on Weekdays", ylab = "Average number of steps", 
     ylim =c(0, 250))
plot(y = dailyact_we, x = names(dailyact_we), type = "l", xlab = "5-Minute Interval", 
     main = "Daily Activity Pattern on Weekends", ylab = "Average number of steps", 
     ylim =c(0, 250))
```
